04_DESeq_sexes

### 2019-12-22 analysis/04_LDA.Rmd
remove old files, working on new lda analysis

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set treatment levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))

# craete variable that will be critical for subset later on
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))

# ceate new variable for hypothesis testing
c.colData <- c.colData %>%
    mutate(hypothesis = fct_recode(treatment,
                            "non-parental" = "control",
                            "non-parental" = "bldg",
                            "incubation" = "lay",
                            "incubation" = "inc.d3",
                            "incubation" = "inc.d9",
                            "incubation" = "inc.d17",
                            "hatchling.care" = "hatch",
                            "hatchling.care" = "n5",
                            "hatchling.care" = "n9"))

# drop columsn that are bad for model AND not needed for later joining
c.colData$bird <- NULL
c.colData$tissue <- NULL
c.colData$study <- NULL
c.colData$group <- NULL
head(c.colData)

# prep col data for subsequent analyses
colDataHyp <- subsetcolData2(c.colData, c("female_hypothalamus", "male_hypothalamus"))
colDataPit <- subsetcolData2(c.colData, c("female_pituitary", "male_pituitary"))
colDataGon <- subsetcolData2(c.colData, c("female_gonad", "male_gonad"))
```

```{r dds}
dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus") )
dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary", "male_pituitary" ) )
dds.gonad <- subsetDESeq2(c.colData,  c.countData, c("female_gonad", "male_gonad") )
```

## total degs

```{r totalDEGs, results = 'hide', eval = F}
DEGs.pituitary <- returntotalDEGs(dds.pituitary)
plottotalDEGs(DEGs.pituitary, "Male & Female Pituitary DEGs")
```

## Variance stabilized data

```{r vsd}
# create variance stbilized DESeq transformation
vsd.hyp <- vst(dds.hypothalamus, blind=FALSE) 
vsd.pit <- vst(dds.pituitary, blind=FALSE) 
vsd.gon <- vst(dds.gonad, blind=FALSE)

# save as a dataframe for downstream analysis
vsd.hyp.df <- vsd.dataframe(vsd.hyp) 
vsd.pit.df <- vsd.dataframe(vsd.pit) 
vsd.gon.df <- vsd.dataframe(vsd.gon) 
```

```{r viewwritevsd}
savevsdfiles(vsd.hyp.df, colDataHyp, "hyp")
savevsdfiles(vsd.pit.df, colDataPit, "pit")
savevsdfiles(vsd.gon.df, colDataGon, "gon")
```

## PCA

```{r pca}
hypPCA <- returnPCAs2(vsd.hyp)
plotPC12(hypPCA, "Hypothalamus")

pitPCA <- returnPCAs2(vsd.pit)
plotPC12(pitPCA, "Pituitary")

gonPCA <- returnPCAs2(vsd.gon)
plotPC12(gonPCA, "Gonad")
```

@@

## plot genes in a  PRL WGCNA module

```{r wgcna, eval = T, fig.width=6, fig.height=6}
# pitutiary expression
plotWGCNAcandidates(vsd.pit, c("NP_990797.2"), colDataPit, "PRL expression, Pituitary")

# COX1,  PRL, 
topPCAloadings <- c("NP_006917.1",  "NP_990797.2")
h <- plotWGCNAcandidates(vsd.hyp, topPCAloadings, colDataHyp, "Hypothalamus")
p <- plotWGCNAcandidates(vsd.pit, topPCAloadings, colDataPit, "Pituitary")

plot_grid(h + theme(axis.text.x = element_blank(), legend.position = "none"),
          p , nrow = 2, rel_heights = c(0.45,0.55))
```

```{r wgcna2, eval = T, fig.width=12, fig.height=8}
PRL_associated <- read.csv("../results/08_PRL_associated.csv", header = T)
prl36 <-  PRL_associated$entrezid[1:57]
prl72 <-  PRL_associated$entrezid[58:102]

plotWGCNAcandidates(vsd.pit, prl36, colDataPit, "A-K genes with PRL pattern in the pituitary")
plotWGCNAcandidates(vsd.pit, prl72, colDataPit, "K-Z genes with PRL pattern in the pituitary")
```

```{r ctsGE}
rts <- assay(vsd.pit)
```

>>>>

```{r readfiles}
colData <- read.csv("../metadata/00_samples.csv", row.names = 1)

characterization <- colData %>% filter(study == "charcterization")
manipluation <- colData %>% filter(study == "manipulation")

vsdshyp <- read.csv("../results/06_hypallvsd.csv")
vsdspit <- read.csv("../results/06_pitallvsd.csv")
vsdsgon <- read.csv("../results/06_gonallvsd.csv")

vsdall <- full_join(vsdshyp, vsdspit)
vsdall <- full_join(vsdall, vsdsgon)
head(vsdall)
```

@@
